---
title: "Homework-03"
format: html
author: "Rebecca-Jane Gullette"
---

https://github.com/rebecca-jane-gullette/Gullette-RebeccaJane_homework-03.git

##Read in Packages

```{r}
# general use
library(tidyverse)
library(readxl)
library(here)
library(janitor)

# visualizing pairs
library(GGally)

# model selection
library(MuMIn)

# model predictions
library(ggeffects)

# model tables
library(gtsummary)
library(flextable)
library(modelsummary)

drought_exp <- read_xlsx(path = here("data", 
                                     "Valliere_etal_EcoApps_Data.xlsx"),
                         sheet = "First Harvest")

# quick look at data 
str(drought_exp)
class(drought_exp)
```

##Cleaning Data 
```{r}
# cleaning
drought_exp_clean <- drought_exp %>% 
  clean_names() %>% # nicer column names
  mutate(species_name = case_when( # adding column with species scientific names
    species == "ENCCAL" ~ "Encelia californica", # bush sunflower
    species == "ESCCAL" ~ "Eschscholzia californica", # California poppy
    species == "PENCEN" ~ "Penstemon centranthifolius", # Scarlet bugler
    species == "GRICAM" ~ "Grindelia camporum", # great valley gumweed
    species == "SALLEU" ~ "Salvia leucophylla", # Purple sage
    species == "STIPUL" ~ "Nasella pulchra", # Purple needlegrass
    species == "LOTSCO" ~ "Acmispon glaber" # deerweed
  )) %>% 
  relocate(species_name, .after = species) %>% # moving species_name column after species
  mutate(water_treatment = case_when( # adding column with full treatment names
    water == "WW" ~ "Well watered",
    water == "DS" ~ "Drought stressed"
  )) %>% 
  relocate(water_treatment, .after = water) # moving water_treatment column after water
```

```{r}
ggpairs(drought_exp_clean, # data frame
        columns = c("leaf_dry_weight_g", # columns to visualize
                    "sla", 
                    "shoot_g", 
                    "root_g", 
                    "total_g"), 
        upper = list(method = "pearson")) + # calculating Pearson correlation coefficient
  theme_bw() + # cleaner theme
  theme(panel.grid = element_blank()) # getting rid of gridlines
```

```{r total biomass differ between species}
ggplot(data = drought_exp_clean, # data frame
       aes(x = reorder(species_name, # reordering x-axis
                       -total_g, # in reverse order of mean total mass
                       fun = mean), # calculating mean to reorder
           y = total_g)) + # y-axis
  geom_jitter(width = 0.1, # narrow jitter
              height = 0) # not jittering points up and down
```

```{r total biomass between water treatment}
ggplot(data = drought_exp_clean, # data frame
       aes(x = water_treatment, # x-axis
           y = total_g)) + # y-axis
  geom_jitter(width = 0.1, # narrow jitter
              height = 0) # not jittering points up and down
```
```{r specific leaf area on total biomass}
ggplot(data = drought_exp_clean, # data frame
       aes(x = sla, # x-axis
           y = total_g)) + # y-axis
  geom_point() # scatterplot
```
## 0. Null model

```{r}
model0 <- lm(total_g ~ 1, # formula
             data = drought_exp_clean) # data frame
```

## 1. total biomass as a function of SLA, water treatment, and species

```{r}
# saturated model
model1 <- lm(total_g ~ sla + water_treatment + species_name,
             data = drought_exp_clean)

par(mfrow = c(2, 2))
plot(model1)
# you might get a warning when you run this code - that is ok!
```
## 2. total biomass as a function of SLA and water treatment

```{r}
model2 <- lm(total_g ~ sla + water_treatment,
             data = drought_exp_clean)

plot(model2)
```

## 3. total biomass as a function of SLA and species

```{r}
model3 <- lm(total_g ~ sla + species_name,
             data = drought_exp_clean)

plot(model3)
```

## 4. total biomass as a function of water treatment and species

```{r}
model4 <- lm(total_g ~ water_treatment + species_name,
             data = drought_exp_clean)

plot(model4)
```

# Model selection

From the `MuMIn` package

```{r}
model.sel(model0,
          model1, 
          model2, 
          model3,
          model4)
```

#Problem 1. Multiple linear regression: model selection and construction

##a. Make a table or list of all the models from class and the last one you constructed on your own. Write a caption for your table.

```{r}
modelsummary::modelsummary(
  list(
    "Null" = model0,
    "Model 1" = model1,
    "Model 2" = model2,
    "Model 3" = model3,
    "Model 4" = model4
  )
) 
```


